\chapter{Introduction}

This chapter covers some general issues related to this manual and the game it
describes.  

\section{What is Call to Power II?}

General info on what kind of game CtP2 is: turn based, strategy, manage resources, fight wars, blah blah.

\section{Outline}

Here we briefly describe the structure of this manual.

Chapter~\vref{chap:gettingStarted} provides a quick-start guide and a cursory
description of the in-game interface to enable you to understand what is going
on as quickly as possible.

Part~\vref{part:gameMechanics} covers all the game mechanics in more depth,
organising them by concept.

Part~\vref{part:furtherTopics} describes the multiplayer game formats, provides
strategy advice for play, explains how to alter the game settings and options
and gives advice for troubleshooting.

Part~\vref{part:Appendicies} includes appendicies giving detailed statistics on
all the units, buildings, advances and other in-game objects.

The manual ends with an index on page~\pageref{chap:index}.

\section{Disclaimer}

This manual is very much a work in progress.  Do not expect it to be
comprehensive as yet.  Furthermore, like any documentation describing something
under active development, it is likely that parts of this manual will be out of
date or otherwise inaccurate.

\section{Legal Notes}

Legal information on the game, source code and manual.

\section{Credits}

Who helped make CtP2, AE and the manual?

\section{Goals}

Explore, grow, build, research, fight, trade, negotiate, win.

\section{Apolyton Edition}
\index{Poly@'Poly}

Explain what the AE is and its history.

\section{Differences from Civilization: Call to Power}

Explain main differences between games maybe???

\section{Differences from Civilization 3}

Explain main differences between games maybe???

